---
title: "SocCult Exam"
author: "Amalie Lysgaard Andersen"
date: "6/5/2019"
output: html_document
---
errythang
```{r setup + packages}
setwd("~/Library/Mobile Documents/com~apple~CloudDocs/4. Aarhus Universitet/4. Semester/2. Social and cultural dynamics/SocCult_Exam")

pacman::p_load("devtools", "rstan", "rethinking", "dplyr", "brms", "tidyverse", "bayesplot", "RColorBrewer", "cowplot", "ggeffects")
```

```{r data}
df <- read.csv("soccult_data_updated.csv")
tom <- read.csv("tom_trials.csv")
```

```{r no NAs df}
#data w/o na's 
new <- df %>%
  select(ID, age, gender, condition, set_solo, perf_solo, perf_pair, perf_diff, competence, amount, steps, payoff, payoff_partner, payoff_diff, win, tom)

```


                                                   ### 1. initial look at data
  
Participants
```{r Participants info}
mean(df$age)
sum(df$gender == "F")
```

Plots
```{r simple: playing around w plots}

ggplot(df) + 
  aes(steps, competence) +
  geom_count(aes(color = tom))

ggplot(df) +
  aes(steps, competence) +
  geom_tile(aes(fill = tom))

ggplot(df) +
  aes(steps, tom) +
  geom_tile(aes(fill = competence)) 

ggplot(df) +
  aes(competence, tom) +
  geom_tile(aes(fill = steps))

#grid
ggplot(df) +
  aes(steps, tom) + 
  geom_point() +
  facet_wrap(~competence)

#simple
one <- ggplot(df) +
  aes(tom, steps) + 
  geom_point() 
two <- ggplot(df) +
  aes(competence, steps) + 
  geom_point() 

plot_grid(one, two)
```

```{r simple initial plots}
ggplot(df2, aes(steps)) +
  geom_histogram(binwidth = 0.7) #+
  #scale_x_continuous(breaks = seq(0, 100, 5))


ggplot(df2, aes(condition, payoff)) + 
  geom_point(aes(colour = condition)) 

ggplot(df, aes(ID, tom)) + 
  geom_point()

ggplot(df, aes(condition, tom)) + 
  geom_point(aes(colour = condition)) 

ggplot(df, aes(payoff, tom)) + 
  geom_point(aes(colour = condition)) 

ggplot(df, aes(condition, win)) +
  geom_histogram()

ggplot(df, aes(steps, condition)) +
  geom_point()


# with models
plot(df$steps ~ df$condition) 
plot(df$steps ~ df$tom) 
plot(df$steps ~ df$competence) 
# ^GRINEREN --> jo lavere comp jo flere steps


# density plots
dens (df$steps) 
dens (df$competence)

hist(df$steps) 
```

```{r spaghettiplot of tom}
display.brewer.all() 

xlab <- "Trial" 
ylab <- "Measure of Theory of Mind"
num <- 22
getPalette <- colorRampPalette(brewer.pal(9,"Accent"))

ggplot(tom, aes(trial, tom)) +
  geom_point(aes(colour = ID), size = 2) +
  scale_color_manual(values = getPalette(num)) +
  geom_line(aes(colour = ID), size = 0.5) +
  labs(x = xlab, y = ylab) +
  scale_y_continuous(breaks = c(-10:20)) +
  scale_x_continuous(breaks = c(1:3)) +
  theme_minimal() +
  labs(title = "Performance across trials in Matching Pennies game")
```

```{r density, raw data on condition}
palpal <- colorRampPalette(brewer.pal(9,"Spectral"))
numb <- 2
xlab2 <- "Steps from $100" 
ylab2 <- "Density"

ggplot(df, aes(steps)) + 
  geom_density(aes(fill = condition, colour = condition), alpha = 0.3) + 
  facet_wrap(~condition) +
  scale_color_manual(values = palpal(numb)) +
  scale_fill_manual(values = palpal(numb)) +
  theme_minimal() +
  labs(x = xlab2, y = ylab2, title = "Distribution of steps across conditions") 
```

 
                                                    ### 2. Priors + modelling

- Riccardo's suggestions for prior: 
priors in model 0

intercept 	norm (50, 10)
beta (mani)	norm (20, 10)		what is the difference between comp and coop
  scep beta	norm (0, 5 or 10)	
sigma		norm (0, 10)		how much is the model wrong

# FIRST ROUND (m1.1 + m1.2)

m0.0 NOTHING 
```{r}
get_prior(steps ~ 1, data = df, family = gaussian())

m0.0 <- brm(
  steps ~ 1 , 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m0.0
#checking divergents
pairs(m0.0, np = nuts_params(model0.0))
#plot
plot(m0.0)
#ppcheck
pp_check(m0.0, nsamples = 50) 
```

m0  condition
```{r brm: model0} 
get_prior(steps ~ 1 + condition, data = df, family = gaussian())
  
m0 <- brm( 
  steps ~ 1 + condition, 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"),  
            set_prior("normal(0, 10)", class = "sigma")),
  data = df,
  sample_prior = "only"
)

# sample_prior: Indicate if samples from all specified proper priors should be drawn additionally to the posterior samples (defaults to "no"). Among others, these samples can be used to calculate Bayes factors for point hypotheses via hypothesis. If set to "only", samples are drawn solely from the priors ignoring the likelihood, which allows among others to generate samples from the prior predictive distribution. In this case, all parameters must have proper priors.

#calling results
m0
#checking divergents
pairs(m0, np = nuts_params(model0))
#plot
plot(m0)
#ppcheck
pp_check(m0, nsample = 50)  #color_scheme_set("red")
```

1.1. condition + tom
```{r brm: model 1.1: tom}
get_prior(steps ~ 1 + condition + tom, data = df, family = gaussian())

m1.1 <- brm(
  steps ~ 1 + condition + tom, 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)
 
#calling results
m1.1
#checking divergents
pairs(m1.1, np = nuts_params(m1.1))
#plot
plot(m1.1)
#ppcheck
pp_check(m1.1)
```

1.2. condition + competence
```{r brm: model 1.2: comp} 
get_prior(steps ~ 1 + condition + competence, data = df, family = gaussian())

m1.2 <- brm(
  steps ~ 1 + condition + competence,  
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m1.2 
#checking divergents
pairs(m1.2, np = nuts_params(m1.2))
#plot
plot(m1.2)
#ppcheck
pp_check(m1.2)
```

1.3. tom + competence
```{r}
get_prior(steps ~ 1 + tom + competence, data = df, family = gaussian()) 
   
m1.3 <- brm(
  steps ~ 1 + tom + competence,  
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
) 

#calling results
m1.3 
#checking divergents
pairs(m1.3, np = nuts_params(m1.3))
#plot
plot(m1.3)
#ppcheck
pp_check(m1.3, nsamples = 50)
```

WAIC comparison: m0 + m0.0
```{r WAIC Round 1}
# compute and save the WAIC information for the next three models 
m0 <- add_criterion(m0, "waic") 
m0.0 <- add_criterion(m0.0, "waic")

# compare the WAIC estimates
w1 <- loo_compare(m0, m0.0, criterion = "waic")
print(w1, simplify = F)

#Calculating weights
model_weights(m0, m1.1, m1.2, m1.3, weights = "waic") %>% 
  round(digits = 2)


### RESULTS
#     elpd_diff se_diff elpd_waic se_elpd_waic p_waic se_p_waic waic   se_waic
#m1.3    0.0       0.0  -105.1       4.2          3.5    1.2     210.3    8.5 
#m1.1   -2.3       1.6  -107.4       4.6          3.5    1.3     214.8    9.3 
#m1.2   -3.0       2.4  -108.2       3.2          3.1    1.0     216.4    6.4 
#m0     -3.3       2.3  -108.4       3.2          2.1    0.8     216.9    6.5 
```

WAIC comparison: m1.1 + m1.2 + m1.3 + m0
```{r WAIC Round 1}
# compute and save the WAIC information for the next three models 
m1.1 <- add_criterion(m1.1, "waic") 
m1.2 <- add_criterion(m1.2, "waic")
m1.3 <- add_criterion(m1.3, "waic")

# compare the WAIC estimates
w1 <- loo_compare(m0, m1.1, m1.2, m1.3, criterion = "waic")
print(w1, simplify = F)

#Calculating weights
model_weights(m0, m1.1, m1.2, m1.3, weights = "waic") %>% 
  round(digits = 2)


### RESULTS
#     elpd_diff se_diff elpd_waic se_elpd_waic p_waic se_p_waic waic   se_waic
#m1.3    0.0       0.0  -105.1       4.2          3.5    1.2     210.3    8.5 
#m1.1   -2.3       1.6  -107.4       4.6          3.5    1.3     214.8    9.3 
#m1.2   -3.0       2.4  -108.2       3.2          3.1    1.0     216.4    6.4 
#m0     -3.3       2.3  -108.4       3.2          2.1    0.8     216.9    6.5 
```


# SECOND ROUND (m1.1 + m2.2)
2.1. condition * tom
```{r brm: model 2.1: tom} 
get_prior(steps ~ 1 + condition * tom, data = df, family = gaussian()) 
 
m2.1 <- brm(
  steps ~ 1 + condition * tom, 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m2.1
#checking divergents
pairs(m2.1, np = nuts_params(m2.1))
#plot
plot(m2.1)
#ppcheck
pp_check(m2.1)
```

WAIC comparison: (m1.1 + m2.1)
```{r WAIC Round 2}
# compute and save the WAIC information for the next three models
m2.1 <- add_criterion(m2.1, "waic") 

# compare the WAIC estimates
w2 <- loo_compare(m1.1, m2.1, criterion = "waic")
print(w2, simplify = F)

#Calculating weights
model_weights(m0, m1.1, m1.2, weights = "waic") %>% 
  round(digits = 2)
```


# THIRD ROUND (m3.1 + m3.2)
3.1. condition * tom + competence
```{r} 
# Getting the priors  
get_prior(steps ~ 1 + condition * tom + competence, df, family = gaussian) 
  
m3.1 <- brm(  
  steps ~ 1 + condition * tom + competence, 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m3.1
#checking divergents
pairs(m3.1, np = nuts_params(m3.1))
#plot
plot(m3.1)
#ppcheck
pp_check(m3.1, nsamples = 50)
```

3.2. condition * tom * competence
```{r}
# Getting the priors 
get_prior(steps ~ 1 + condition * tom * competence, df, family = gaussian)
 
m3.2 <- brm(
  steps ~ 1 + condition * tom * competence, 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results 
m3.2
#checking divergents
pairs(m3.2, np = nuts_params(m3.2))
#plot
plot(m3.2)
#ppcheck
pp_check(m3.2)
```

WAIC COMPARISON MODEL 2.1 + 3.1
```{r WAIC Round 3}
# compute and save the WAIC information for the next three models
m3.1 <- add_criterion(m3.1, "waic")  
m3.2 <- add_criterion(m3.2, "waic") 

# compare the WAIC estimates
w <- loo_compare(m3.1, m3.2, criterion = "waic")
print(w, simplify = F)
#Calculating weights
model_weights(m3.1, m3.1, weights = "waic") %>% 
  round(digits = 2)  
```


# FOURTH [ALL OR NOTHING] ROUND (m4.0, m4.1, m4.2, m4.3, m4.4)
4.0. tom * competence
```{r}
get_prior(steps ~ 1 + tom * competence, data = df, family = gaussian())
 
m4.0 <- brm(
  steps ~ 1 + tom * competence,  
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)
 
#calling results
m4.0 
#checking divergents
pairs(m4.0, np = nuts_params(m4.0))
#plot
plot(m4.0)
#ppcheck
pp_check(m4.0)
```

4.1. condition + tom + competence
```{r}  
# Getting the priors  
get_prior(steps ~ 1 + condition + tom + competence, df, family = gaussian)
  
m4.1 <- brm(
  steps ~ 1 + condition + tom + competence, 
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m4.1
#checking divergents
pairs(m4.1, np = nuts_params(m4.1))
#plot
plot(m4.1)
#ppcheck
pp_check(m4.1, nsamples = 50)
```

4.2. condition + tom * competence
```{r}  
get_prior(steps ~ 1 + condition + tom * competence, data = df, family = gaussian())
 
m4.2 <- brm( 
  steps ~ 1 + condition + tom * competence,  
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m4.2 
#checking divergents
pairs(m4.2, np = nuts_params(m4.2))
#plot
plot(m4.2)
#ppcheck
pp_check(m4.2)
```

4.3. tom
```{r}
get_prior(steps ~ 1 + tom , data = df, family = gaussian())
 
m4.3 <- brm(
  steps ~ 1 + tom,  
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

#calling results
m4.3
#checking divergents
pairs(m4.3, np = nuts_params(m4.3))
#plot
plot(m4.3)
#ppcheck
pp_check(m4.3)
```

4.4 competence
```{r}
get_prior(steps ~ 1 + competence , data = df, family = gaussian())

m4.4 <- brm( 
  steps ~ 1 + competence,  
  prior = c(set_prior("normal(20, 5)", class = "Intercept"), 
            set_prior("normal(0, 10)", class = "b"), 
            set_prior("normal(0, 10)", class = "sigma")),
  data = df
)

 #calling results
m4.4
#checking divergents
pairs(m4.4, np = nuts_params(m4.4))
#plot 
plot(m4.4)
#ppcheck
pp_check(m4.4)
```


# WAIC TIME
```{r adding waic criterion for all models}
m0 <- add_criterion(m0, "waic") 
m0.0 <- add_criterion(m0.0, "waic")
m1.1 <- add_criterion(m1.1, "waic") 
m1.2 <- add_criterion(m1.2, "waic")
m1.3 <- add_criterion(m1.3, "waic")
m2.1 <- add_criterion(m2.1, "waic")
m3.1 <- add_criterion(m3.1, "waic")
m3.2 <- add_criterion(m3.2, "waic")
m4.0 <- add_criterion(m4.0, "waic") 
m4.1 <- add_criterion(m4.1, "waic")
m4.2 <- add_criterion(m4.2, "waic") 
m4.3 <- add_criterion(m4.3, "waic") 
m4.4 <- add_criterion(m4.4, "waic") 
```

```{r WAIC Round 4}
# compare the WAIC estimates
w <- loo_compare(m4.1, m4.2, criterion = "waic")
print(w, simplify = F)

#Calculating weights
model_weights(m3.1, m3.1, weights = "waic") %>% 
  round(digits = 2)  
```

```{r WAIC alround tester}
#all waic  
wall <- loo_compare(m1.1, m1.2, m1.3, m2.1, m3.1, m3.2, m4.1, m4.2, m4.3, m4.4, m4.0, m0, m0.0, criterion = "waic")
print(wall, round = 2) 
 
wyas <- loo_compare(m0.0, m0, criterion = "waic")
print(wyas, simplify = F)

wtest <- loo_compare(m4.1, m4.11, criterion = "waic")
print(wtest) 

wwehu <- loo_compare(m4.1, m0.0, m0, criterion = "waic")
print(wwehu, round = 2) 
 
wwow <- loo_compare(m4.1, m1.3, m0, m3.1, m0.0, criterion = "waic") 
print(wwow, round = 2) 
```





                                                 ### 3. Final plots


# Different plot 
s
```{r marginal effects all plots}
brms::marginal_effects(m3.1)
brms::marginal_effects(m3.2)
brms::marginal_effects(m4.1)
brms::marginal_effects(m4.11)
brms::marginal_effects(m1.3)
```

```{r plots of m1.3}
### pred vs actual
#creating predictions from model
pred <- predict(m1.3) 
pred <- as.data.frame(pred)
pred$actual <- df$steps
pred$condition <- df$condition

#plot
ggplot(pred, aes(actual, Estimate)) +
  geom_smooth(color = "#449999", method = "lm", se = F) +
  geom_point(color = "Dark red") +
  geom_abline(xintercept = 100, yintercept = 100, linetype = "dashed") +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_minimal() +
  labs(title = "Observation vs. Prediction - Model 3", x = "Observed values (steps)", y = "Model predictions (steps)")

### just simple marginal
this <- brms::marginal_effects(m1.3)
plot(this, plot = F)[[2]] +
  scale_color_brewer(type = "div") +
  #scale_fill_brewer(type = "qual", palette = "Set1") +
  #labs(title = "Marginal effects of ToM & steps", x = "ToM", y = "Steps from $100") +
  theme_minimal()

```

```{r plots of m3.1}
#creating predictions from model
pred2 <- predict(m3.1) 
pred2 <- as.data.frame(pred2)
pred2$actual <- df$steps
pred2$condition <- df$condition

#as factor
#pred$diagnosis <- as.factor(pred$diagnosis)

#plots
ggplot(pred2) +
  aes(condition, Estimate) +
  geom_violin()

plot1 <- ggplot(pred2) +
  aes(condition, Estimate) +
  geom_boxplot() 
plot2 <- ggplot(pred2) +
  aes(condition, actual) +
  geom_boxplot()  

plot_grid(plot1, plot2)

# marginal, m3.1
that <- brms::marginal_effects(m3.1)
plot(that, plot = F)[[4]] +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  labs(title = "Marginal effects of ToM & condition", x = "ToM", y = "Steps from $100") +
  theme_minimal()

```

```{r trying out kld}
pacman::p_load("LaplacesDemon", "NetworkToolbox") 

pred <- predict(m1.3)
pred <- as.data.frame(pred)
pred$actual <- df$steps
pred$condition <- df$condition

b <- KLD(pred$actual, pred$Estimate)
b

kld(pred$actual, pred$Estimate)
```

```{r competence + steps faceted by condition}
ggplot(df, aes(tom, steps, color = condition)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "Steps by tom")

num2 <- 2
getPalette <- colorRampPalette(brewer.pal(2,"Set1"))

ggplot(df, aes(condition, steps, color = condition)) +
  geom_point() +
  scale_color_manual(values = getPalette(num2)) +
  geom_smooth(method = "lm") +
  labs(title = "Steps by condition")


ggplot(df, aes(competence, steps, color = condition)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "Steps by competence")
```

